---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE, 
  warning = FALSE,
  error = FALSE
)
```
# deepSentimentR

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip) [![License: GPL-3](https://img.shields.io/badge/License-GPLv3-yellow.svg)](https://opensource.org/licenses/GPL-3.0) [![Travis build status](https://travis-ci.org/adityamangal410/deepSentimentR.svg?branch=master)](https://travis-ci.org/adityamangal410/deepSentimentR) [![Coverage status](https://codecov.io/gh/adityamangal410/deepSentimentR/branch/master/graph/badge.svg)](https://codecov.io/github/adityamangal410/deepSentimentR?branch=master)
<!-- badges: end -->

The goal of deepSentimentR is to provide functionality to query, view and analyze sentiment datasets. 

## Installation

<!--
You can install the released version of deepSentimentR from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("deepSentimentR")
```
-->

### Prerequisites

``` r
install.packages("dplyr")
install.packages("keras")
keras::install_keras()
```

### DeepSentimentR

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("adityamangal410/deepSentimentR")
```

## Setup using Rocker

You can use [rocker](https://www.rocker-project.org/), which provides a set of docker images with R and different flavours of tidyverse, ML, geospatial etc. 

### Setup using rocker/ml

```
docker run -d --name rstudio -p 8787:8787 -e PASSWORD=<PASSWORD> -e USER=<USER> -v /Users/<USER>/Desktop:/Desktop rocker/ml
```

Go to http://localhost:8787 to login to the rstudio server.

``` r
devtools::install_github("adityamangal410/deepSentimentR")
```

### Setup using rocker/tidyverse  

Experimental steps  

```
docker run -d --name rstudio -p 8787:8787 -e PASSWORD=<PASSWORD> -e USER=<USER> -v /Users/<USER>/Desktop:/Desktop rocker/tidyverse
```

Go to http://localhost:8787 to login to the rstudio server.

``` r
library(tidyverse)
system("python -V") # make sure you're using python3
#If Not, login to docker container and run following
#rm /usr/bin/python
#ln -s /usr/bin/python3 /usr/bin/python
#apt-get install curl
system("curl https://bootstrap.pypa.io/get-pip.py -o get-pip.py")
system("python get-pip.py")
system("pip3 install virtualenv")
install.packages("keras")
keras::install_keras()
devtools::install_github("adityamangal410/deepSentimentR")
```


## Example

Visualize polarity frequency based on given filters:  

```{r example}
library(dplyr)
library(deepSentimentR)
data("sentiment140_train")
sentiment140_train %>% 
  freq_by_polarity(user_list = c("", "mcraddictal"), 
                   start_date_time = lubridate::as_datetime("2009-05-30"),
                   end_date_time = lubridate::as_datetime("2009-06-03"),
                   keyword_list = c(""))
```

Use pre-trained LSTM model using Glove Embeddings to make polarity predictions for the sample test dataset:  

```{r example2}
predict_polarity_keras(model_load_path = system.file("extdata",
                                                     "train_glove_lstm.rds",
                                                     package = "deepSentimentR",
                                                     mustWork = TRUE))
```


Check out full reference and examples at - https://adityamangal410.github.io/deepSentimentR/
